Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-22 implementation.tex
Added text about buffering, benchmark
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1135 ccdd05d5-56c1-461a-bf98-85f6876523b4

>>>>

\gstreamer\ pipelines typically operate with very low (in some consumer applications, imperceptibly low) latency rather than in true real-time because signals are partitioned into blocks of samples, or \emph{buffers}.  This affords a number of advantages, including amortizing the overhead of passing signals between elements and grouping together sequences of similar operations.  However, buffering a signal incurs a latency of up to one buffer length.  This latency can be made small at the cost of some additional overhead by making the buffers sufficiently small.

### 2011-06-22 results.tex
minimum match -> minimal match
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1112 ccdd05d5-56c1-461a-bf98-85f6876523b4
Added text about buffering, benchmark
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1135 ccdd05d5-56c1-461a-bf98-85f6876523b4

The \SNR\ loss is to be compared with the mismatch of 0.03 that arises from the
discreteness of template bank designed for a minimum match of 0.97.  We will consider
an acceptable target \SNR\ loss to be a factor of 10 smaller than this, that is, no more
than 0.003.

@@

>>>>

The \SNR\ loss is to be compared with the mismatch of 0.03 that arises from the
discreteness of template bank designed for a minimal match of 0.97.  We will consider
an acceptable target \SNR\ loss to be a factor of 10 smaller than this, that is, no more
than 0.003.

@@

\subsection{Measured latency and overhead}

Our \gstreamer\ pipeline for measuring impulse responses contained instrumentation that would not be necessary for an actual search, including additional interpolation filters to bring the early-warning outputs back to the full sample rate and additional outputs for recording signals to disk.

We wrote a second, stripped pipeline to evaluate the actual latency and computational overhead.  We executed this pipeline on one of the submit machines of the \LIGO-Caltech cluster, a Sun Microsystems Sun Fire\texttrademark\ X4600~M2 server with eight quad-core 2.7~GHz AMD Opteron\texttrademark\ 8384 processors.  This test consumed 78\% of the capacity of just one out of the 32 cores, maintaining a constant latency of 5~s.

The measured overhead is consistent to within an order of magnitude with the lower bound from the \flops\ budget.  The additional overhead is probably dominated by thread synchronization.  A carefully optimized \gstreamer\ pipeline or a hand-tuned C implementation of the pipeline might reduce overhead further.

The 5~s latency is probably due to buffering and synchronization.  The latency could probably be reduced by carefully tuning buffer lengths at every stage in the pipeline.
